$p$-adic precision

### 2014-02-13 padicprec.tex
I introduced the notion of first order lattice.

Then, there exists a positive real number $\delta$ such that, for all $r 
\in (0, \delta)$, the equality
\begin{equation}
\label{eq:main}
f(v_0 + H) = f(v_0) + f'(v_0) \cdot H.
\end{equation}
holds for each lattice $H$ such that $B^-_E(r) \subset H \subset
B_E(r)$.
\end{lem}

\begin{rmk}
Given a lattice $H_0$ in $E$, the equality \eqref{eq:main} holds for $H 
= \lambda H_0$ with $\lambda$ small enough. To prove this, it is enough 
to apply Lemma \ref{lem:main} to the map $f \circ \iota_{H_0}$ where 
$\iota_{H_0} : (E, \Vert \cdot \Vert_{H_0}) \to (E, \Vert \cdot \Vert)$ 
is the ``identity map'' (which is an homeomorphism by Lemma 
\ref{lem:reductionball}).
\end{rmk}

@@

Let $r \in (0, \delta)$. Let $B^-_E(r) \subset H \subset
B_E(r).$
Let us show that $f$ maps surjectively $H$ onto $f'(0) \cdot H$.

We first prove that $f(H) \subset f'(0) \cdot H$.
Let $x \in H$. By differentiability at $0$, $\Vert f(x)-f'(0) \cdot x \Vert \leq 
\varepsilon \Vert x \Vert $. Therefore, if $y=f(x)-f'(0) \cdot x,$ then $\Vert y \Vert \leq \varepsilon r$. As a consequence of the open mapping Theorem, $B_F(\varepsilon r) \subset 
\varepsilon C  f'(0) \cdot B_E(r),$ and there exists $x_1 \in \varepsilon C  \cdot B_E(r)$ such that $f'(0) \cdot x_1 =y.$ Since $\varepsilon C <1,$  $x_1 \in B^-_E(r) \subset H.$ and finally, $f(x)= f'(0) \cdot (x-x_1) \in f'(0) \cdot H$.

We shall now prove surjectivity. Let $y \in f'(0) \cdot H$. 
Let $x_0 \in H$ be such that $y = f'(0) \cdot x_0$. Since $f(x_0) \in f'(0) \cdot H,$ we can define $z_0 \in E$ to be such that $f'(0) \cdot z_0 = y - f(x_0)$ and $\Vert z_0 \Vert \leq C \cdot \Vert y - f(x_0) \Vert$.
We then define by induction two sequences $(x_n)$ and $(z_n)$ by
\begin{itemize}
\item $x_{n+1}=x_n+z_n$ and 
\item $z_n \in E$ is such that
$f'(0) \cdot z_n = y - f(x_n)$ and
$\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$.
\end{itemize}
By definition, $\Vert z_0 \Vert < C \Vert y_0\Vert$ and $y 
= -f(x_0)+f'(0) \cdot x_0$. Therefore, $\Vert y \Vert < \varepsilon 
\Vert x_0 \Vert \leq \varepsilon r$. Thus $\Vert z_0 \Vert \leq 
\varepsilon C \cdot r < r$. Thus, $z_0$ and $x_1=x_0+z_0$  lie in $B^-_E(r) 
\subset H$. An easy induction involving an analoguous computation
shows that the sequences $(x_n)$ and $(z_n)$ are well defined and
take their values in $H$.

@@

\todofor{Tristan}{Improve the end of this Subsection}

@@

@@

The theory of finite dimensional $K$-manifolds is presented for example 
in \cite{Schneider}, parts 8 and 9. In this Section, we shall work with
a slightly different notion of manifolds which allows also Banach vector 
spaces of infinite dimension.
More precisely, for us, a \emph{ differentiable $K$-manifold} 
(or just \emph{$K$-manifold} for short) is the data of a topological 
space $V$ together with an open covering $V = \bigcup_{i \in I} V_i$ 
(where $I$ is some set) and, for all $i \in I$, an homeomorphism 
$\varphi_i : V_i \to U_i$ where $U_i$ is an open subset of a $K$-Banach 
$E_i$ such that for all $i,j \in I$ for which $V_i \cap V_j$ is 
nonempty, the composite map
\begin{equation}
\label{eq:psiij}
\psi_{ij} : 
\varphi_i(V_{ij}) \stackrel{\varphi_i^{-1}}{\longrightarrow} 
V_{ij} \stackrel{\varphi_j}{\longrightarrow} \varphi_j(V_{ij})
\quad \text{(with } V_{ij} = V_i \cap V_j \text{)}
\end{equation}
is  differentiable at all points in its domain. We recall that
the mappings $\varphi_i$ above are the so-called \emph{charts} and that
their collection defines an \emph{atlas} of $V$. In the sequel, we shall
assume further that the open covering $V = \bigcup_{i \in I} V_i$ is
locally finite, which means that every point $x \in V$ lies only in a
finite number of $V_i$'s. Trivial examples of $K$-manifolds are 
$K$-Banach themselves.

@@

\todofor{David/Xavier}{Complete this Subsection}

@@

\begin{deftn}
Let $x \in V$. 
A \emph{precision datum} at $x$ is a lattice in the tangent space 
$T_x V$.
\end{deftn}

\begin{deftn}
Let $x \in V$ and $H$ be a precision datum at $x$.
For all $i \in I$ such that $x \in V_i$, we set:
$$x + O_i(H) = \varphi_i^{-1}\big(\varphi_i(x) + \varphi_i'(x)(H)\big) 
\subset V.$$
\end{deftn}

We emphasize that if $x$ belongs to $V_i$ and $V_j$ at the same time,
the subsets $x + O_i(H)$ and $x + O_j(H)$ differ in general. It is 
however not the case if $H$ is small enough.

\begin{prop}
\label{prop:independence}
Let $x \in V$ and $H$ be a precision data at $x$.
If $H$ is ``small enough'' \todo{(precise!)}, the subset $x + 
O_i(H)$ does not depend on $i$.
\end{prop}

\begin{proof}
Let $i$ and $j$ be two indices such that $x$ belongs to $V_i$ and $V_j$. 
Set $x_i = \varphi_i(x) \in E_i$ and $H_i = \varphi'_i(x)(H)$. Going 
back to the definitions, we see that the equality $x + O_i(H) = x + 
O_j(H)$ is equivalent to:
$$\psi_{ij}(x_i + H_i) = \psi_{ij}(x_i) + \psi'_{ij}(x_i)(H_i)$$ 
where we recall that $\psi_{ij}$ is defined by Eq.~\eqref{eq:psiij}.
\todo{Complete the proof and write the correct condition on $H$!}
\end{proof}

Under the assumption of Proposition \ref{prop:independence}, we define
$x + O(H)$ as $x + O_i(H)$ for some (equivalently all) $i$ such that
$x \in V_i$.

@@

In order to restrict ourselves to elements $x$ lying in a dense 
computable subset, we need to compare $x_0 + O(H_0)$ with varying $x + 
O(H)$ when $x$ and $x_0$ are close enough.
Let us first examine the situation in a fixed given chart: we fix some 
index $i \in I$ and pick two elements $x_0$ and $x$ in $V_i$. We consider 
in addition a precision datum $H_0$ at $x_0$ and we want to produce a 
precision datum $H$ at $x$ such that $x_0 + O_i(H_0) = x + O_i(H)$. 
We remark that the tangent spaces $T_{x_0} V$ and $T_x V$ are both 
isomorphic to $E_i$ via the maps $\varphi'_i(x_0)$ and 
$\varphi'_i(x)$ respectively. A natural candidate for $H$ is then:
\begin{equation}
\label{eq:Hprime}
H = \left(\varphi'_i(x)^{-1} \circ \varphi'_i(x_0)\right) (H_0).
\end{equation}
And indeed, one can check that $x + O(H)$ is indeed equal to $x_0 + 
O(H_0)$ as soon as $x$ and $x_0$ are close enough in the following sense: 
the difference $\varphi_i(x) - \varphi_i(x_0)$ lies in the lattice $H_i = 
\varphi'_i(x_0)(H_0)$. We furthermore have a property of independence on 
$i$.

\begin{prop}
We assume that $H_0$ is ``small enough'' \todo{(precise!)} and that $x$ is close 
enough to $x_0$. Then
\begin{enumerate}[(i)]
\item the lattice $H$ defined by \eqref{eq:Hprime} does not depend 
on $i$,
\item the pairs $(x,H)$ and $(x_0, H_0)$ satisfy the assumptions of
Proposition \ref{prop:independence} and we have $x + O(H) = x_0 + O(H_0)$.
\end{enumerate}
\end{prop}

\begin{proof}
We first prove (i). For an index $i$ such that $x, x_0 \in V_i$, let us 
denote by $f_i : T_{x_0} V \to T_x V$ the composite $\varphi'_i(x)^{-1} 
\circ \varphi'_i(x_0)$. Given an extra index $j$ satisfying the same
assumption, we know by \todo{(find a reference)} that the difference
$f_i - f_j$ goes to $0$ when $x$ converges to $x_0$. Since $H_0$ is open
in $T_{x_0} V$, this implies that $(f_j - f_i)(H_0)$ contains $f_i(H_0)$ and 
$f_j(H_0)$ if $x$ and $x_0$ are close enough. Now, pick $w \in f_j(H_0)$ and 
write it $w = f_j(v)$ with $v \in H_0$. Then $w$ is equal to $f_i(v) + 
(f_j - f_i)(v)$ and then belongs to $f_i(H_0)$ because each summand does. 
Therefore $f_j(H_0) \subset f_i(H_0)$. The inverse inclusion can be proved 
in the same way.

We now prove (ii). \todo{Write the proof.}
\end{proof}

@@

\todofor{David/Xavier}{Extend Lemma \ref{lem:main} to manifolds.}

>>>>

Then, there exists a positive real number $\delta$ such that, for all $r 
\in (0, \delta)$, the equality
\begin{equation}
\label{eq:main}
f(v_0 + H) = f(v_0) + f'(v_0) (H).
\end{equation}
holds for each lattice $H$ such that $B^-_E(r) \subset H \subset
B_E(r)$.
\end{lem}

\begin{rmk}
\label{rmk:mainlemma}
Given a lattice $H_0$ in $E$, the equality \eqref{eq:main} holds for $H 
= \lambda H_0$ with $\lambda$ small enough. To prove this, it is enough 
to apply Lemma \ref{lem:main} to the map $f \circ \iota_{H_0}$ where 
$\iota_{H_0} : (E, \Vert \cdot \Vert_{H_0}) \to (E, \Vert \cdot \Vert)$ 
is the ``identity map'' (which is an homeomorphism by Lemma 
\ref{lem:reductionball}).
\end{rmk}

@@

Let $r \in (0, \delta)$. Let $B^-_E(r) \subset H \subset
B_E(r).$
Let us show that $f$ maps surjectively $H$ onto $f'(0) (H)$.

We first prove that $f(H) \subset f'(0) (H)$.
Let $x \in H$. By differentiability at $0$, $\Vert f(x)-f'(0) \cdot x \Vert \leq 
\varepsilon \Vert x \Vert $. Therefore, if $y=f(x)-f'(0) \cdot x,$ then $\Vert y \Vert \leq \varepsilon r$. As a consequence of the open mapping Theorem, $B_F(\varepsilon r) \subset 
\varepsilon C  f'(0) \cdot B_E(r),$ and there exists $x_1 \in \varepsilon C  \cdot B_E(r)$ such that $f'(0) \cdot x_1 =y.$ Since $\varepsilon C <1,$  $x_1 \in B^-_E(r) \subset H.$ and finally, $f(x)= f'(0) \cdot (x-x_1) \in f'(0) (H)$.

We shall now prove surjectivity. Let $y \in f'(0) (H)$.
Let $x_0 \in H$ be such that $y = f'(0) \cdot x_0$. Since $f(x_0) \in f'(0) (H)$, we can define $z_0 \in E$ to be such that $f'(0) \cdot z_0 = y - f(x_0)$ and $\Vert z_0 \Vert \leq C \cdot \Vert y - f(x_0) \Vert$.
We then define by induction two sequences $(x_n)$ and $(z_n)$ by
\begin{itemize}
\item $x_{n+1}=x_n+z_n$ and 
\item $z_n \in E$ is such that
$f'(0) \cdot z_n = y - f(x_n)$ and
$\Vert z_n \Vert \leq C \cdot \Vert y - f(x_n) \Vert$.
\end{itemize}
By definition, $\Vert z_0 \Vert < C \Vert y_0\Vert$ and $y 
= -f(x_0)+f'(0) \cdot x_0$. Therefore, $\Vert y \Vert < \varepsilon 
\Vert x_0 \Vert \leq \varepsilon r$. Thus $\Vert z_0 \Vert \leq 
\varepsilon C \cdot r < r$. Thus, $z_0$ and $x_1=x_0+z_0$  lie in $B^-_E(r) 
\subset H$. An easy induction involving an analoguous computation
shows that the sequences $(x_n)$ and $(z_n)$ are well defined and
take their values in $H$.

@@

We introduce the following definition:

\begin{deftn}
\label{def:firstorder}
Let $E$ and $F$ be two $K$-Banach and $f : U \rightarrow F$ be a 
function defined on an open subset $U$ of $E$. Let also $v_0$ be a 
point in $U$.
A lattice $H$ in $E$ is called a \emph{first order lattice} for $f$ at 
$v_0$ if
$$f(v_0 + H) = f(v_0) + f'(v_0) (H).$$
\end{deftn}

By Lemma \ref{lem:main} and Remark \ref{rmk:mainlemma}, if $f$ is 
differentiable at $v_0$ and $f'(v_0)$ is surjective, then for all 
lattice $H$ in $E$, there exists a constant $\delta$ such that $aH$ 
is a first order lattice for $f$ at $v_0$ for all $a \in B_E(r)$.

\subsection{The case of locally analytic functions}

\todofor{Tristan}{Improve this Subsection}

@@

Using Lemma \ref{lem:reductionball}, one can extend Proposition 
\ref{prop:locanalytic} to arbitrary lattices $H$ as follows. Given $H$
a lattice in some $K$-Banach space $E$, we first define the two
following quantities:
\begin{itemize}
\item $\max(H)$ is the smallest $r$ such that $H \subset B_E(r)$, and
\item $\min(H)$ is the greatest $r$ such that $B_E(r) \subset H$.
\end{itemize}

\begin{cor}
Let $U$ are $V$ be open subsets in $K$-Banach spaces $E$ and $F$ 
respectively and $f : U \to V$ be a locally analytic function. Let 
also $v_0$ be a point in $U$ such that $f'(v_0)$ is surjective. 
Then a lattice $H$ in $E$ is a first order lattice for $f$ at $v_0$ as 
soon as $\max(H)^2 < C_{f, v_0} \cdot \min(H)$ where $C_{f, v_0}$ is a 
constant depending only on $f$ and $v_0$ which can be determined 
explicitely from the Newton polygon of $f$ at $v_0$.
\end{cor}

\begin{proof}
We apply Proposition \ref{prop:locanalytic} to the composite $f \circ 
\iota_H$ where $\iota_H : (E, \Vert \cdot \Vert_H) \to (E, \Vert \cdot 
\Vert)$ is induced by identity map.
\end{proof}

@@

The theory of finite dimensional $K$-manifolds is presented for example 
in \cite{Schneider}, parts 8 and 9. In this Section, we shall work with
a slightly different notion of manifolds which allows also Banach vector 
spaces of infinite dimension.
More precisely, for us, a \emph{ differentiable $K$-manifold} 
(or just \emph{$K$-manifold} for short) is the data of a topological 
space $V$ together with an open covering $V = \bigcup_{i \in I} V_i$ 
(where $I$ is some set) and, for all $i \in I$, an homeomorphism 
$\varphi_i : V_i \to U_i$ where $U_i$ is an open subset of a $K$-Banach 
$E_i$ such that for all $i,j \in I$ for which $V_i \cap V_j$ is 
nonempty, the composite map
\begin{equation}
\label{eq:psiij}
\psi_{ij} : 
\varphi_i(V_{ij}) \stackrel{\varphi_i^{-1}}{\longrightarrow} 
V_{ij} \stackrel{\varphi_j}{\longrightarrow} \varphi_j(V_{ij})
\quad \text{(with } V_{ij} = V_i \cap V_j \text{)}
\end{equation}
is differentiable at all points in its domain. We recall that
the mappings $\varphi_i$ above are the so-called \emph{charts}. The
$\psi_{ij}$'s are the transition maps. The collection of $\varphi_i$'s
and $\psi_{ij}$'s is called an \emph{atlas} of $V$. In the sequel, we 
shall assume further that the open covering $V = \bigcup_{i \in I} V_i$ 
is locally finite, which means that every point $x \in V$ lies only in a 
finite number of $V_i$'s. Trivial examples of $K$-manifolds are 
$K$-Banach themselves.

@@

@@

\begin{deftn}
Let $x \in V$.
A \emph{precision datum} at $x$ is a lattice in the tangent space 
$T_x V$ such that for all indices $i$ and $j$ with $x \in U_i \cap
U_j$, the image of $T_x V$ in $E_i$ is a first order lattice for
$\psi_{ij}$ at $\varphi_i(x)$ (\emph{cf} Definition \ref{def:firstorder}.
\end{deftn}

\begin{rmk}
The definition of a precision datum at $x$ depends not only on $x$
and the manifold $V$ where it lies but also on the chosen atlas that
defines $V$.
\end{rmk}

\begin{prop}
\label{prop:independence}
Let $x \in V$ and $H$ be a precision datum at $x$.
The subset:
$$\varphi_i^{-1}\big(\varphi_i(x) + \varphi_i'(x)(H)\big) 
\subset V$$
does not depend on the index $i$ such that $x \in V_i$.
\end{prop}

\begin{proof}
Let $i$ and $j$ be two indices such that $x$ belongs to $V_i$ and $V_j$. 
Set $x_i = \varphi_i(x) \in E_i$ and $H_i = \varphi'_i(x)(H)$. The 
equality 
$$\varphi_i^{-1}\big(\varphi_i(x) + \varphi_i'(x)(H)\big) 
\subset V = \varphi_j^{-1}\big(\varphi_j(x) + \varphi_j'(x)(H)\big) 
\subset V$$
is clearly equivalent to
$\psi_{ij}(x_i + H_i) = \psi_{ij}(x_i) + \psi'_{ij}(x_i)(H_i)$.
and the latter holds because $H_i$ is a first order lattice
of $\psi_{ij}$ at the point $x_i$ by definition.
\end{proof}

We are now in position to define $x + O(H)$.

\begin{deftn}
Let $x \in V$ and $H$ be a precision datum at $x$. We set:
$$x + O(H) = \varphi_i^{-1}\big(\varphi_i(x) + \varphi_i'(x)(H)\big)
\subset V$$
for some (equivalenty, all) $i$ such that $x \in V_i$.
\end{deftn}

@@

In order to restrict ourselves to elements $x$ lying in a dense 
computable subset, we need to compare $x_0 + O(H_0)$ with varying $x + 
O(H)$ when $x$ and $x_0$ are close enough. Let us first examine the 
situation in a fixed given chart: we fix some index $i \in I$ and pick 
two elements $x_0$ and $x$ in $V_i$. We consider in addition a lattice 
$\tilde H_0$ in $E_i$ --- which should be think as $\varphi'_i(x_0) 
(H_0)$ --- and we want to produce a lattice $\tilde H$ such that 
$\varphi_i(x_0) + \tilde H_0 = \varphi_i(x) + \tilde H$. Of course 
$\tilde H = \tilde H_0$ does the job as soon as 
$\varphi_i(x) - \varphi_i(x_0) \in \tilde H_0$. 
Now, we remark that the tangent spaces $T_{x_0} V$ and $T_x V$ are both 
isomorphic to $E_i$ via the maps $\varphi'_i(x_0)$ and $\varphi'_i(x)$ 
respectively. A natural candidate for $H$ is then:
\begin{equation}
\label{eq:Hprime}
H = \left(\varphi'_i(x)^{-1} \circ \varphi'_i(x_0)\right) (H_0).
\end{equation}
With this choice, $x + O(H) = x_0 + O(H_0)$ as soon as $x$ and $x_0$ are 
close enough in the following sense: the difference $\varphi_i(x) - 
\varphi_i(x_0)$ lies in the lattice $\varphi'_i(x_0)(H_0)$. We 
furthermore have a property of independence on $i$.

\begin{prop}
Let $x_0 \in V$ and $H_0$ be a precision datum at $x_0$.
Then, for all $x$ sufficiently close to $x_0$,
\begin{enumerate}[(i)]
\item the lattice $H$ defined by \eqref{eq:Hprime} does not depend 
on $i$ and is a precision datum at $x$, and
\item we have $x + O(H) = x_0 + O(H_0)$.
\end{enumerate}
\end{prop}

\begin{proof}
We first prove (i). For an index $i$ such that $x, x_0 \in V_i$, let us 
denote by $f_i : T_{x_0} V \to T_x V$ the composite $\varphi'_i(x)^{-1} 
\circ \varphi'_i(x_0)$. Given an extra index $j$ satisfying the same
assumption, we know by \todo{(find a reference)} that the difference
$f_i - f_j$ goes to $0$ when $x$ converges to $x_0$. Since $H_0$ is open
in $T_{x_0} V$, this implies that $(f_j - f_i)(H_0)$ contains $f_i(H_0)$ and 
$f_j(H_0)$ if $x$ and $x_0$ are close enough. Now, pick $w \in f_j(H_0)$ and 
write it $w = f_j(v)$ with $v \in H_0$. Then $w$ is equal to $f_i(v) + 
(f_j - f_i)(v)$ and then belongs to $f_i(H_0)$ because each summand does. 
Therefore $f_j(H_0) \subset f_i(H_0)$. The inverse inclusion can be proved 
in the same way. The fact that $H$ is a precision datum at $x$ is easy
and left to the reader.
Finally, if $x$ is close enough to $x_0$, we can check (ii) in the 
charts but this was already done.
\end{proof}

@@

\todofor{Xavier}{Extend Lemma \ref{lem:main} to manifolds.}

\subsection{Examples}

\todofor{David}{Elliptic curves and grassmannians}
